The SMT Competition 2015--2018

### 2019-04-13 2018/report/SMT-COMP.tex
update tables

Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, 30,717 non-incremental benchmarks
from the SMT Library were not used for the competition because their
status was unknown.  However, this approach was criticized for
imitating existing solvers rather than truly innovating.  In~2016,
the SMT Library contained 29,724 non-incremental benchmarks with
unknown status; solver performance on these was evaluated but reported
separately.  Since 2017, all non-incremental benchmarks are eligible
for the competition's Main track regardless of status, and they
directly affect the competition results.  Solver responses for such
benchmarks are assumed to be correct, unless there is a disagreement
between two or more (otherwise sound) solvers.  Benchmarks with
disagreements are removed from the competition results.  In 2016,
there were disagreements on 79 benchmarks; in 2017, only one benchmark
was removed; and in 2018, there were no disagreements between
otherwise sound solvers.

For incremental benchmarks, the situation is different: an incremental
benchmark is ineligible for the Application track if its first {\tt
  check-sat} command has unknown status.  Otherwise, the prefix of the
benchmark that only contains {\tt check-sat} commands with known
status is eligible.  The main reason for this difference is that the
competition tools for the Application track (in particular, the trace
executor) have not been adapted yet to support {\tt check-sat}
commands with unknown status.

@@

@@

\begin{table}
\caption{Alternative winners if only Satisfiable (top) or Unsatisfiable (bottom) benchmarks had been considered. These results are in reference to those given in Table~\ref{tab:winners-main}.\label{tab:results:sat:unsat}}
\centering
\begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.25em]}c}
\toprule
Division         &  2015                 &  2016                    &  2017                  &  2018                      \\
\hline \hline
AUFLIA  & CVC3  &   &   & Alt-Ergo  \\
AUFNIRA & \nc{Z3}  & CVC4 \nc{Z3}  & CVC4 \nc{Z3}  & \nc{Z3}  \\
NIA     & CVC3  &   &   &   \\
NRA     & CVC3  & CVC4  & CVC4  & CVC4\nc{Z3}Z3  \\
QF\_AUFBV        & Yices  & Yices2  &   & Yices2  \\
QF\_LIA  & \nc{Z3}  &   &   &   \\
QF\_UFIDL        & \nc{Z3}  & \nc{Z3}  & \nc{Z3}  & \nc{Z3} veriT  \\
QF\_UFNIA        & CVC3  &   & CVC4  &   \\
UFLIA   & \nc{Z3}  &   &   &   \\
UFLRA   & CVC3  & CVC4  &   &   \\
AUFLIRA &   & CVC4  & CVC4  & Alt-Ergo  \\
BV      &   & \nc{Z3}  & Boolector  & Boolector  \\
QF\_ABV  &   & Yices2  & Yices2  &   \\
QF\_BV   &   & MinkeyRink  &   &   \\
QF\_NIA  &   & SMT-RAT  &   &   \\
UF      &   & Vampire  &   &   \\
UFBV    &   & Boolector  &   &   \\
UFNIA   &   & CVC4  & CVC4  & CVC4  \\
LRA     &   &   & veriT+Redlog  &   \\
QF\_LRA  &   &   & Yices2  & Yices2  \\
QF\_NIRA &   &   & CVC4  &   \\
QF\_NRA  &   &   & \nc{Z3}  &   \\
UFDTLIA &   &   & CVC4  &   \\
ALIA    &   &   &   & Alt-Ergo  \\
QF\_FP   &   &   &   & CVC4  \\
QF\_UFBV &   &   &   & Yices2  \\
\bottomrule
\end{tabular}

\begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.25em]}c}
\toprule
Division         &  2015                 &  2016                    &  2017                  &  2018                      \\
\hline \hline
QF\_AUFNIA       & \nc{Z3}  & \nc{Z3}  &   &   \\
QF\_IDL  & CVC4  &   & \nc{Z3}  & CVC4 \nc{Z3}  \\
QF\_LIA  & SMTInterpol  & SMTInterpol  & SMTInterpol  & SMTInterpol  \\
QF\_NIA  & CVC3  &   & Yices2  &   \\
BV      &   & CVC4  &   &   \\
NIA     &   & CVC4  & Vampire  & Vampire  \\
QF\_UFNRA        &   & CVC4 \nc{Z3}  & CVC4  & CVC4 \nc{Z3}  \\
UFIDL   &   & Vampire \nc{Z3}  & \nc{Z3}  &   \\
QF\_AUFBV        &   &   & CVC4  &   \\
QF\_BV   &   &   & Boolector  &   \\
QF\_UFLIA        &   &   & \nc{Z3}  &   \\
UF      &   &   & CVC4  & CVC4  \\
UFLRA   &   &   & veriT  & veriT  \\
QF\_NRA  &   &   &   & CVC4  \\
\bottomrule
\end{tabular}

@@

The time limits used in each year are relatively long (20 or 40 minutes) whereas some applications of SMT solvers (e.g. software verification) make use of very short time limits. We ask the question \emph{how would the results change if we consider a much shorter time limit of 24 seconds?}\footnote{This time limit was inspired by a conversation with Liana Hadarean as it is a time limit being used in one application of SMT solvers in Amazon.}. The results of this analysis are given in Table~\ref{tab:results:24s}. As in the previous, this highlights the different winners under this new scheme only. Here we see STP winning a division for the first time in our analysis (QF\_BV in 2016) and Yices2 performing well.

@@

\begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.25em]}c}
\toprule
Division         &  2015                 &  2016                    &  2017                  &  2018                      \\
\hline \hline
AUFLIRA & CVC3  &   &   & Alt-Ergo  \\
AUFNIRA & \nc{Z3}  &   &   & \nc{Z3}  \\
QF\_BV   & CVC4  & STP \nc{Z3}  &   & Boolector  \\
QF\_LIA  & Yices  & Yices2  & Yices2  &   \\
QF\_LRA  & Yices  & Yices2  & Yices2  & Yices2  \\
QF\_UF   & veriT  & veriT  &   &   \\
UFNIA   & \nc{Z3}  &   &   & \nc{Z3}  \\
AUFLIA  &   & Vampire \nc{Z3}  & veriT  &   \\
QF\_ABV  &   & Yices2  &   &   \\
QF\_ANIA &   & \nc{Z3}  & \nc{Z3}  &   \\
QF\_UFNRA        &   & CVC4 \nc{Z3}  & CVC4  & \nc{Z3}  \\
UF      &   & Vampire  &   &   \\
UFBV    &   & Boolector  &   &   \\
UFIDL   &   & Vampire \nc{Z3}  & \nc{Z3}  &   \\
QF\_NIA  &   &   & Yices2  & Yices2  \\
QF\_NIRA &   &   & Yices2 Z3  & Yices2 \nc{Z3}  \\
QF\_NRA  &   &   & \nc{Z3}  &   \\
UFLRA   &   &   & veriT  & veriT  \\
BV      &   &   &   & Boolector \nc{Z3}  \\
NRA     &   &   &   & CVC4 \nc{Z3}  \\
UFLIA   &   &   &   & \nc{Z3}  \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Do Families Matter?}

@@

Whilst producing these results we noticed that the description of what constitutes a family had been incorrectly interpreted by the scoring scripts in 2016-2018. After fixing this misinterpretation we note that only a single division winner changes. As this is in the past we are choosing not to retroactively change the winner (and all tables in this paper reflect the previously published results) but the scripts have been fixed for future iterations of the competition.

@@

>>>>

Another paradigm shift concerns the treatment of benchmarks with
unknown status.  Because of the difficulty to judge solver responses
for such benchmarks, SMT Competitions until 2015 only used benchmarks
whose status was known (i.e., that had been solved previously by
existing SMT solvers).  In 2015, 30,717 non-incremental benchmarks
from the SMT Library were not used for the competition because their
status was unknown.  However, this approach was criticized for
imitating existing solvers rather than truly innovating.  In~2016,
the SMT Library contained 29,724 non-incremental benchmarks with
unknown status; solver performance on these was evaluated but reported
separately.  Since 2017, all non-incremental benchmarks are eligible
for the competition's Main track regardless of status, and they
directly affect the competition results.  Solver responses for such
benchmarks are assumed to be correct, unless there is a disagreement
between two or more (otherwise sound e.g. on problems with known status) solvers.  Benchmarks with
disagreements are removed from the competition results.  This conservative approach was chosen over a voting system as this situation is expected to be rare as the SMT-LIB organisers go to a significant effort prior to each competition to update the statuses of problems. A manual investigation is therefore preferred over an automated resolution. Disagreements are reported to the solvers involved and in all cases have led to bugs in solvers being identified and fixed. 
In 2016,
there were disagreements on 79 benchmarks; in 2017, only one benchmark
was removed; and in 2018, there were no disagreements between
otherwise sound solvers.

For incremental benchmarks, the situation is different: an incremental
benchmark is ineligible for the Application track if its first {\tt
  check-sat} command has unknown status.  Otherwise, the prefix of the
benchmark that only contains {\tt check-sat} commands with known
status is eligible.  The main reason for this difference is that the
competition tools for the Application track (in particular, the trace
executor) have not been adapted yet to support {\tt check-sat}
commands with unknown status. Even though the trace executor has now been modified, we cannot check what impact this may have had on previous results without rerunning all tracks as the trace executor drives the solver.

@@

The use of StarExec ensures the \emph{reproducibility} of the competition. The publicly accessible space \url{https://www.starexec.org/starexec/secure/explore/spaces.jsp?id=2641} stores all of the information required to rerun the competition e.g. all benchmarks and copies of all tools. In addition, a GitHub organisation (\url{https://github.com/SMT-COMP/smt-comp}) has been setup to archive previous results and store the tools used by organisers to run the competition.

@@

\begin{table}
\caption{Alternative winners if only Satisfiable (top) or Unsatisfiable (bottom) benchmarks had been considered. These results are in reference to those given in Table~\ref{tab:winners-main}.\label{tab:results:sat:unsat}}
\centering
\begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.25em]}c}
\toprule
Division         &  2015                 &  2016                    &  2017                  &  2018                      \\
\hline \hline
ALIA    &   &   &   & Alt-Ergo  \\
AUFLIA  & CVC3  &   &   & Alt-Ergo  \\
AUFLIRA &   & CVC4  & CVC4  & Alt-Ergo  \\
AUFNIRA & \nc{Z3}  & CVC4 \nc{Z3}  & CVC4 \nc{Z3}  & \nc{Z3}  \\
BV      &   & \nc{Z3}  & Boolector  & Boolector  \\
LRA     &   &   & veriT+Redlog  &   \\
NIA     & CVC3  &   &   &   \\
NRA     & CVC3  & CVC4  & CVC4  & CVC4 \nc{Z3}  \\
QF\_ABV  &   & Yices2  & Yices2  &   \\
QF\_AUFBV        & Yices  & Yices2  &   & Yices2  \\
QF\_BV   &   & \nc{Z3}/MinkeyRink  &   &   \\
QF\_FP   &   &   &   & CVC4  \\
QF\_LIA  & \nc{Z3}  &   &   &   \\
QF\_LRA  &   &   & Yices2  & Yices2  \\
QF\_NIA  &   & SMT-RAT  &   &   \\
QF\_NIRA &   &   & CVC4  &   \\
QF\_NRA  &   &   & \nc{Z3}  &   \\
QF\_UFBV &   &   &   & Yices2  \\
QF\_UFIDL        & \nc{Z3}  & \nc{Z3}  & \nc{Z3}  & \nc{Z3} veriT  \\
QF\_UFNIA        & CVC3  &   & CVC4  &   \\
UF      &   & Vampire  &   & Vampire/-  \\
UFBV    &   & Boolector  &   &   \\
UFDTLIA &   &   & CVC4  &   \\
UFLIA   & \nc{Z3}  &   &   &   \\
UFLRA   & CVC3  & CVC4  &   &   \\
UFNIA   &   & CVC4  & CVC4  & CVC4  \\
\bottomrule
\end{tabular}

\begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.25em]}c}
\toprule
Division         &  2015                 &  2016                    &  2017                  &  2018                      \\
\hline \hline
BV      &   & CVC4  &   &   \\
NIA     &   & CVC4  & Vampire  & Vampire  \\
QF\_AUFBV        &   &   & CVC4  &   \\
QF\_AUFNIA       & \nc{Z3}  & \nc{Z3}  &   &   \\
QF\_BV   &   &   & Boolector  &   \\
QF\_IDL  & CVC4  &   & \nc{Z3}  & CVC4 \nc{Z3}  \\
QF\_LIA  & SMTInterpol  & SMTInterpol  & SMTInterpol  & SMTInterpol  \\
QF\_NIA  & CVC3  &   & Yices2  &   \\
QF\_NRA  &   &   &   & CVC4  \\
QF\_UFLIA        &   &   & \nc{Z3}  &   \\
QF\_UFNRA        &   & CVC4 \nc{Z3}  & CVC4  & CVC4 \nc{Z3}  \\
UF      &   &   & CVC4  & -/CVC4  \\
UFIDL   &   & Vampire \nc{Z3}  & \nc{Z3}  &   \\
UFLRA   &   &   & veriT  & veriT  \\
\bottomrule
\end{tabular}

@@

The time limits used in each year are relatively long (20 or 40 minutes) whereas some applications of SMT solvers (e.g. software verification) make use of very short time limits. We ask the question \emph{how would the Main track results change if we consider a much shorter time limit of 24 seconds?}\footnote{This time limit was inspired by a conversation with Liana Hadarean as it is a time limit being used in one application of SMT solvers in Amazon.}. We only consider the Main track here as this is the setting where short time limits are most likely to be relevant. The results of this analysis are given in Table~\ref{tab:results:24s}. As in the previous, this highlights the different winners under this new scheme only. Here we see STP winning a division for the first time in our analysis (QF\_BV in 2016) and Yices2 performing well.

@@

\begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.25em]}c}
\toprule
Division         &  2015                 &  2016                    &  2017                  &  2018                      \\
\hline \hline
AUFLIA  &   & Vampire \nc{Z3}  & veriT  &   \\
AUFLIRA & CVC3  &   &   & -/Alt-Ergo  \\
AUFNIRA & \nc{Z3}  &   &   & \nc{Z3}  \\
BV      &   &   &   & Boolector/Boolector \nc{Z3}  \\
NRA     &   &   &   & CVC4/CVC4 \nc{Z3}  \\
QF\_ABV  &   & Yices2  &   &   \\
QF\_ANIA &   & \nc{Z3}  & \nc{Z3}  &   \\
QF\_BV   & CVC4  & STP \nc{Z3}  & Boolector+CaDiCaL/-  & Boolector  \\
QF\_LIA  & Yices  & Yices2  & Yices2  &   \\
QF\_LRA  & Yices  & Yices2  & Yices2  & Yices2  \\
QF\_NIA  &   &   & Yices2  & Yices2  \\
QF\_NIRA &   &   & Yices2 \nc{Z3}  & Yices2 \nc{Z3}  \\
QF\_NRA  &   &   & \nc{Z3}  &   \\
QF\_UF   & veriT  & veriT  &   &   \\
QF\_UFNRA        &   & CVC4 \nc{Z3}  & CVC4  & \nc{Z3}  \\
UF      &   & Vampire  &   &  Vampire/- \\
UFBV    &   & Boolector  &   &   \\
UFIDL   &   & Vampire \nc{Z3}  & \nc{Z3}  &   \\
UFLIA   &   &   &   & \nc{Z3}  \\
UFLRA   &   &   & veriT  & veriT  \\
UFNIA   & \nc{Z3}  &   &   & CVC4/\nc{Z3}  \\
\bottomrule
\end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%
\subsection{Do Families Matter?}

@@

Whilst producing these results we noticed that the description of what constitutes a family had been incorrectly interpreted by the scoring scripts in 2016-2018. After fixing this misinterpretation we note that only a single division winner changes - in 2017 AUFNIRA should have been won by

. As this is in the past we are choosing not to retroactively change the winner (and all tables in this paper reflect the previously published results) but the scripts have been fixed for future iterations of the competition.

@@

%%%%%%%%%%%%%%%%%
\subsection{What Impact Did Adding Unknowns Have?}

In 2017 the Main track ran with benchmarks with unknown status (removing any that had disagreements). An interesting question is to what extent this affected the winners in the Main track. Table~\ref{tab:results:unknowns} gives the alternative winners if these benchmarks had not been included.

\begin{table}
\caption{Alternative winners if benchmarks with unknown status had not bee included in 2017 onwards.\label{tab:results:unknowns}}
\centering
\begin{tabular}{r@{\hskip 1em}>{\columncolor[gray]{.95}[.25em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][.5em]}c@{\hskip 1em}>{\columncolor[gray]{.95}[.5em][0.25em]}c}
\toprule
Division                       &  2017                  &  2018                      \\
\hline\hline
AUFNIRA & CVC4 & \\
QF\_NIRA & \nc{Z3} & CVC4 \nc{Z3} \\
QF\_NRA & \nc{Z3} & \\
UFDTLIA & CVC4 & \\
BV & & Boolector \\
QF\_FP & & CVC4 \nc{Z3} \\
UFNIA & & CVC4 \\
NRA & & -/\nc{Z3} \\
UF & & -/CVC4\\
\bottomrule
\end{tabular}
\end{table}
